Modal Logics in the Coq Proof Assistant

### 2014-12-09 Papers/InProgress/ModalLogicCoq/ModalCoq.tex
Wrap paragraphs

Modal logics extend the usual formal logic language with modal operators ($\Box$ and $\Diamond$) and are characterized by the \emph{necessitation rule}, according to which $\Box A$ is a theorem if $A$ is a theorem, although $A \imp \Box A$ is not necessarily a theorem. Various informal concepts, such as \emph{necessity and possibility}, \emph{knowledge and belief}, \emph and \emph{temporal globality and eventuality}, have been formalized with the help of modal operators. Therefore, modal logics \cite{ModalLogic} are well-established for a wide variety of application domains, ranging from philosophy \cite{ToDo} to formal software verification \cite{ToDo:TLAPlus}.

@@

However, one may wonder whether the embedding approach is adequate
also for \emph{interactive} reasoning, when the user proves theorems
by interacting with a proof assistant. The main goal of this paper is
to strudy this question. Our answer is positive. The \Coq proof
assistant was chosen because of the authors' greater familiarity with
the tactic language of this system.

@@

The simpler ontological argument of Anselm has been automatically verified with \texttt{PVS} by Rushby \cite{Rushby} and with first-order theorem by Oppenheimer and Zalta \cite{Zalta}. Gödel's argument was automatically verified in our previous work on embedding-based fully automated modal theorem proving \cite{ArXiv,AFP}. But this paper presents the first fully interactive and detailed formalization of this proof in a proof assistant.

>>>>

Modal logics extend the usual formal logic language with modal
operators ($\Box$ and $\Diamond$) and are characterized by the
\emph{necessitation rule}, according to which $\Box A$ is a theorem if
$A$ is a theorem, although $A \imp \Box A$ is not necessarily a
theorem. Various informal concepts, such as \emph{necessity and
possibility}, \emph{knowledge and belief}, \emph and \emph{temporal
globality and eventuality}, have been formalized with the help of
modal operators. Therefore, modal logics \cite{ModalLogic} are well-
established for a wide variety of application domains, ranging from
philosophy \cite{ToDo} to formal software verification
\cite{ToDo:TLAPlus}.

@@

However, one may wonder whether the embedding approach is adequate
also for \emph{interactive} reasoning, when the user proves theorems
by interacting with a proof assistant. The main goal of this paper is
to study this question. Our answer is positive. The \Coq proof
assistant was chosen because of the authors' greater familiarity with
the tactic language of this system.

@@

The simpler ontological argument of Anselm has been automatically
verified with \texttt{PVS} by Rushby \cite{Rushby} and with first-
order theorem provers by Oppenheimer and Zalta \cite{Zalta}. Gödel's
argument was automatically verified in our previous work on embedding-
based fully automated modal theorem proving \cite{ArXiv,AFP}. But this
paper presents the first fully interactive and detailed formalization
of this proof in a proof assistant.
